---
title: "Anlysis"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE)
```

```{r, results='hide'}
library("Seurat")
library("monocle")
library("DT")
project_dir <- "C:/Users/Yiwei Zhao/Desktop/BRC/Analysis/modi_honey/"
project_names <- c("1_Tg_plus_neg", "2_Tg_plus_pos", "3_Tg_Tg_neg", "4_Tg_Tg_pos")
merged <- readRDS(paste0(project_dir,"merged.rds"))
```

## Overview of Data
The table below summarizes the expression of Ins1, Ins2 and GFP.

Ins1 is expressed in all cells in all 4 samples. The mean expression of Ins1 in tg_tg_neg sample is much lower than in the other 3.

Ins2 is epressed in all cells in the heterozygous samples. It is expressed in some cells in the homozygous samples, but the level is much lower.

GFP is expressed in most cells in GFP-pos samples and in a few cells in GFP-neg samples.

```{r get gene stats}
get_gene_stats <- function(cells, gene, data.use){
  gene_pos_cells <- cells[data.use[gene,cells]>0]
  stats <- c(length(gene_pos_cells),
             median(data.use[gene,cells]),
             mean(data.use[gene,cells]),
             min(data.use[gene,cells]), 
             max(data.use[gene,cells])
             )
  names(stats) <- c(paste0(gene,"_num_cells"),
                    paste0(gene,"_median"),
                    paste0(gene,"_mean"),
                    paste0(gene,"_min"),
                    paste0(gene,"_max")
  )
  return(stats)
}

table <- NULL
genes <- c("Ins1","Ins2","SSM2GFP")
for (project_name in project_names){
  cells <- merged@cell.names[merged@meta.data$orig.ident == project_name]
  row <- c(Total_num_cells = length(cells))
  for (gene in genes){
    row <- c(row, get_gene_stats(cells, gene, merged@raw.data))
  }
  
  table <- rbind(table,row)
  # print(c(length(cells),length(Ins1_pos_cells), length(Ins2_pos_cells), length(gfp_pos_cells)))
}
rownames(table) <- project_names
# print(table)
knitr::kable(table)
# merged@data <- log(merged@raw.data+1)
```

The y-axis in the violin plot below is log normalized count. It seems that in each of the GFP-neg samples, there is a dinstinct group of cells that have very high expression of Ins1 (close to the level in the GFP-pos samples). 

```{r}
VlnPlot(merged, features.plot = genes, group.by = "orig.ident",x.lab.rot = T)
```

Based on the overview of the data, we plan to do the following things first:

* run DE analysis on GFP(+) vs GFP(-) in each of the 2 homozygous samples. As for Ins1, since it is expressed in all cells, what criteria do you think we should use to subset Ins1(-) samples?

* The expression of GFP in tg_tg_pos seems to be somewhat uniformly distributed across a range.Therefore, we can use monocle to find genes that are differentially expressed across the range (though this is usually performed on changes across time).

```{r}
# cells <- merged@cell.names[merged@meta.data$orig.ident == project_names[]]
# plot(x = merged@data["Ins1",cells], y = merged@data["Ins2",cells])
```


```{r}
# s3 <- readRDS(sprintf("%s/%s/%s", project_dir, project_names[3],paste0(project_names[3],".raw.rds")))
# s3_cds <- importCDS(s3)
# s3_cds <- estimateSizeFactors(s3_cds)
# s3_cds <- estimateDispersions(s3_cds)

```


## DE analysis of tg_plus_pos vs tg_tg_pos using Seurat
Genes with positive log_fc values are upregulated in tg_plus_pos.
```{r DE 2vs4}
merged <- SetAllIdent(merged, "orig.ident")
markers <- FindMarkers(object = merged,ident.1 = "2_Tg_plus_pos", ident.2 = "4_Tg_Tg_pos")
datatable(markers, filter = "top")
```

